Operational Feature Selection in Gaussian Mixture Models

### 2016-09-27 paper.tex
update

@@

    The results obtained with two different datasets are presented. The first dataset is an airborne hyperspectral dataset and second is a remote sensing dataset made of heterogeneous spatial features. Training set has been created with an equal number of samples for each class and additionally a spatial stratification has been performed. Several size of training set has been tested. For Aisa dataset, experiments have been conducted using 250, 500 and 1000 samples by class and for the Potsdam dataset, 1000 and 50000 samples by class.

@@

    \subsection{Aisa dataset}
    \label{sec:aisa}
    The Aisa dataset has been acquired by the AISA Eagle sensor during a flight campaign over Heves, Hungary. It contains 252 bands ranging from 395 to 975 nm. 16 classes have been defined for a total of 361,971 referenced pixels. Figure~\ref{fig:aisa} shows a colored composition of the multispectral image and the shapefile representing groundtruth and Table \ref{tab:aisa} presents the repartition of the various classes.

@@

    This second dataset is built from a dataset of remote sensing images distributed by the International Society for Photogrammetry and Remote Sensing (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}. The dataset is composed of aerial images of the urban area of Potsdam. The area is cut into 38 patches of 6000x6000 pixels with a resolution of 5cm by pixel and 4 channels are available: Red, Blue, Green and Infrared (RGBIR). A Digital Surface Model with same resolution is also provided and a so-called normalized DSM representing the height above ground. The groundtruth for 24 tiles are provided with 6 classes: Low vegetation, High vegetation, Impervious surfaces, Buildings, Cars, Clutter. Table~\ref{tab:potsdam} summarizes the number of samples of each class in the two images used during experimentations. Figure~\ref{fig:potsdam-dataset} shows the complete area and Figure~\ref{fig:potsdam-expl} a particular patch.

    \begin{figure}[!t]
        \centering
        \includegraphics[width=0.7\columnwidth]{Fig/potsdam_full.png}
        \caption{Full Potsdam dataset.\label{fig:potsdam-dataset}}
    \end{figure}

    \begin{figure}[!t]
        \centering
        \includegraphics[width=0.8\columnwidth]{Fig/potsdam_imgexpl.png}
        \caption{Example of patch with {\bfseries{(a)}} orthophoto, {\bfseries{(b)}} DSM and {\bfseries{(c)}} groundtruth.\label{fig:potsdam-expl}}
    \end{figure}

    \begin{table}[!t]
        \centering
        \caption{Repartition of classes in Aisa dataset.\label{tab:potsdam}}
        \begin{tabular}[b]{lrr}\hline
          Class & Number of samples & Number of samples \\
                & in 5\_11          & in 5\_12 \\
          \hline
          Clutter             & 1,078,611  & 812,038 \\
          Trees               & 4,493,295  & 2,132,368 \\
          Cars                & 900,076    & 1,101,541 \\
          Buildings           & 13,469,575 & 17,501,421 \\
          Low vegetation      & 4,718,219  & 3,210,596 \\
          Impervious surfaces & 11,340,224 & 11,242,036 \\
          \hline
        \end{tabular}
    \end{table}

    In the experiment, the following features are computed using the RGBIR images similar to \cite{tuia2015multiclass}:
    \begin{itemize}
        \item Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2 (15 features) \footnote{\url{https://www.orfeo-toolbox.org//Applications/RadiometricIndices.html}};
        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)\cite{tuia2015multiclass};
        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)\cite{tuia2015multiclass};
        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)\cite{tuia2015multiclass};
        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features)\cite{tuia2015multiclass}.
    \end{itemize}
    The normalized DSM and the raw RGBIR image are added to these 191 features and then then stacked to create a new image with 196 bands. The training is made with tile 5\_11 and test with tile 5\_12. Table~\ref{tab:potsdam-otbsimu} presents the results.

>>>>

\section{Datasets}
\label{sec:datasets}
Experimentation has been conducted on two different datasets. The first dataset, called Aisa, is an airborne hyperspectral dataset and second, called Potsdam, is a remote sensing dataset made of heterogeneous spatial features.

    \subsection{Aisa dataset}
    \label{sec:aisa-dataset}
    The Aisa dataset has been acquired by the AISA Eagle sensor during a flight campaign over Heves, Hungary. It contains 252 bands ranging from 395 to 975 nm. 16 classes have been defined for a total of 361,971 referenced pixels. Figure~\ref{fig:aisa} shows a colored composition of the multispectral image and the shapefile representing groundtruth and Table \ref{tab:aisa} presents the repartition of the various classes.

    \subsection{Potsdam dataset}
    \label{sec:pots-dataset}

    This second dataset is built from a dataset of remote sensing images distributed by the International Society for Photogrammetry and Remote Sensing (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}. The dataset is composed of aerial images of the urban area of Potsdam. The area is cut into 38 patches of 6000x6000 pixels with a resolution of 5cm by pixel and 4 channels are available: Red, Blue, Green and Infrared (RGBIR). A Digital Surface Model with same resolution is also provided and a so-called normalized DSM representing the height above ground. The groundtruth for 24 tiles are provided with 6 classes: Low vegetation, High vegetation, Impervious surfaces, Buildings, Cars, Clutter. Table~\ref{tab:potsdam} summarizes the number of samples of each class in the three images used during experimentations and Figure~\ref{fig:potsdam-expl} the data itself.

    \begin{figure}[!t]
        \centering
        \begin{tabular}{ccc}
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_11_RGB.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_12_RGB.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_3_10_RGB.jpg} \\
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_11_label.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_12_label.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_3_10_label.jpg} \\
            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_05_11_normalized_lastools.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_05_12_normalized_lastools.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_03_10_normalized_lastools.jpg} \\
            {\bfseries{(a)}} & {\bfseries{(b)}}  & {\bfseries{(c)}}\\
        \end{tabular}
        \caption{From top to bottom, image RGB, reference and normalized DSM of: {\bfseries{(a)}} tile 5\_11, {\bfseries{(b)}} tile 5\_12 and {\bfseries{(c)}} tile 3\_10.\label{fig:potsdam-expl}}
    \end{figure}

    \begin{table}[!t]
        \centering
        \caption{Repartition of classes in Aisa dataset.\label{tab:potsdam}}
        \begin{tabular}[b]{lrrr}\hline
          Class & Number of samples & Number of samples & Number of samples \\
                & in 5\_11          & in 5\_12          & in 3\_10 \\
          \hline
          Clutter             & 1,078,611  & 812,038    & 1,890,467 \\
          Trees               & 4,493,295  & 2,132,368  & 8,780,245 \\
          Cars                & 900,076    & 1,101,541  & 434,615 \\
          Buildings           & 13,469,575 & 17,501,421 & 5,128,149 \\
          Low vegetation      & 4,718,219  & 3,210,596  & 11,428,326 \\
          Impervious surfaces & 11,340,224 & 11,242,036 & 8,338,198 \\
          \hline
        \end{tabular}
    \end{table}

    In the experiment, the following features are computed using the RGBIR images similar to \cite{tuia2015multiclass}:
    \begin{itemize}
        \item Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2 (15 features) \footnote{\url{https://www.orfeo-toolbox.org//Applications/RadiometricIndices.html}};
        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)\cite{tuia2015multiclass};
        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)\cite{tuia2015multiclass};
        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)\cite{tuia2015multiclass};
        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features)\cite{tuia2015multiclass}.
    \end{itemize}
    The normalized DSM and the raw RGBIR image are added to these 191 features and then then stacked to create a new image with 196 bands. The training is made with tile 5\_11 and test with tile 5\_12.

@@

    Training set has been created with an equal number of samples for each class and additionally a spatial stratification has been performed. Several size of training set has been tested. For Aisa dataset, experiments have been conducted using 250, 500 and 1000 samples by class and for the Potsdam dataset, 1000 and 50000 samples by class.

@@

    \subsection{Aisa dataset}
    \label{sec:aisa}

@@

    Table~\ref{tab:potsdam-otbsimu} presents the results.
